Local to global N input modelling

### 2025-06-18 manuscript.qmd
Updates for species_layers and background_nitrogen

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate")

@@

* Hazard: Nutrient input. Present as a change between feeds (%?), relative to background levels (ratio?)
  * Possibly ask Camille how the quotas for Tasmanian salmon are calculated
  * Background levels are from [oceanhealthindex.org](https://knb.ecoinformatics.org/view/resource_map_doi:10.5063/F1610XPS) - double-check that it includes wastewater input AND agriculture - should include both - if they're not combined ask Ben because he has them combined somewhere
  * If this is too uncertain perhaps combine into categorical (oligotrophic/eutrophic) classification?
* Exposure: Species presence. Use IUCN/Aquamap data to estimate the % of each species range is occupied by farms
  * give each farm a 1km radius, check for overlaps
  * Gage already has the data to process the Aquamap data
* Vulnerability: species vulnerability to eutrophication + nutrient input - from @butt_trait-based_2022. Possibly also use IUCN threat levels for prioritisation?

@@

```{r uneaten feed analysis}
uneaten_total_stats <- uneaten_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(uneat_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         max = max %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         mean = mean %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))

uneaten_total_compare <- uneaten_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(uneat_biom_mean)) %>% 
  pivot_wider(names_from = "feed", values_from = "total", id_cols = "farm_ID") %>% 
  mutate(diff_past = past-reference,
         diff_future = future-reference,
         diffperc_past = diff_past/reference,
         diffperc_future = diff_future/reference) %>% 
  dplyr::select(-c(reference, future, past)) %>% 
  pivot_longer(names_to = c("measure", "feed"), names_sep = "_", values_to = "value",
               cols = contains("diff"), names_transform = list(measure = as.factor, feed = as.factor)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(value),
          max = maxna(value),
          mean = meanna(value)) %>% 
  mutate(
    min = case_when(measure == "diffperc" ~ 100*min,
                    T ~ min %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1") %>% drop_units()),
    max = case_when(measure == "diffperc" ~ 100*max,
                    T ~ max %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1") %>% drop_units()),
    mean = case_when(measure == "diffperc" ~ 100*mean,
                     T ~ mean %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1") %>% drop_units()),
    )
```

Total uneaten feed across a cohort (548 days) fed the reference feed ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "reference"]` to 
`r uneaten_total_stats$max[uneaten_total_stats$feed == "reference"]` g kg$^{-1}$ fish, with a mean of 
`r uneaten_total_stats$mean[uneaten_total_stats$feed == "reference"]` g kg$^{-1}$ fish. 
Differences between the feeds were very small, but slightly less feed was left uneaten in the reference scenario (@fig-total-uneaten-perday). Uneaten feed in the past scenario ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "past"]` to 
`r uneaten_total_stats$max[uneaten_total_stats$feed == "past"]` g kg$^{-1}$ fish (mean of 
`r uneaten_total_stats$mean[uneaten_total_stats$feed == "past"]` g kg$^{-1}$ fish) while uneaten feed in the future scenario ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "future"]` to 
`r uneaten_total_stats$max[uneaten_total_stats$feed == "future"]` g kg$^{-1}$ (mean of 
`r uneaten_total_stats$mean[uneaten_total_stats$feed == "future"]`  g kg$^{-1}$ fish).

```{r fig-total-uneaten-perday}
#| fig-cap: |
#|   Feed left uneaten when fish are fed the reference (blue), past (red) or future (green) feed. Countries are arranged by descending mean latitude.

uneaten_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(uneat_biom_mean = uneat_biom_mean %>% set_units("g g_fish-1 d-1") %>% set_units("g kg_fish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = uneat_biom_mean, fill = feed)) +
  geom_boxplot() +
  prettyplot() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Uneaten feed", x = "Country")
```

```{r uneaten composition}
#| eval: false

uneaten_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_uneat_stat"]]}) %>% bind_rows()
uneaten_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_uneat_stat"]]}) %>% bind_rows()
uneaten_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_uneat_stat"]]}) %>% bind_rows()

uneaten_all <- uneaten_total %>% 
  select(-contains("biom")) %>% 
  merge(uneaten_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd")))
```

The composition of the uneaten feed is (obviously) the same as the total feed composition detailed in (@sec-feeds).

## Excreted faeces

```{r excreted faeces data}
#| code-summary: Load and prep all excretion data

# The following reads in all the total excretion data and converts the mean and sd values into values/biomass
excreted_total <- farmrun_files %>% 
  purrr::map(function(f) {qs::qread(f)[["total_excr_stat"]]}) %>% 
  bind_rows() %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(excr_mean = mean.x, biomass_mean = mean.y,
         excr_sd = sd.x, biomass_sd = sd.y) %>% 
  mutate(excr_sd = excr_sd/excr_mean,
         biomass_sd = biomass_sd/biomass_mean,
         excr_biom_mean = excr_mean/biomass_mean,
         excr_biom_sd = sqrt(excr_sd^2+biomass_sd^2)*excr_biom_mean)

excreted_total_stats <- excreted_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(excr_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         max = max %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         mean = mean %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))
```

### Amount

Overall, total excretion was lowest when when fish were fed the past feed and highest when fish were fed the reference feed (@fig-total-excreted-perday). 
Total excreted faeces across a cohort (548 days) in the reference scenario ranged from `r excreted_total_stats$min[excreted_total_stats$feed == "reference"]` to 
`r excreted_total_stats$max[excreted_total_stats$feed == "reference"]` g kg$^{-1}$ fish d$^{-1}$, with a mean of 
`r excreted_total_stats$mean[excreted_total_stats$feed == "reference"]`.
Total excreted faeces across a cohort in the past scenario ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "past"]` to 
`r excreted_total_stats$max[excreted_total_stats$feed == "past"]` g kg$^{-1}$ fish d$^{-1}$ (mean of 
`r excreted_total_stats$mean[excreted_total_stats$feed == "past"]` g kg$^{-1}$ fish d$^{-1}$).
Total excreted faeces across a cohort in the future scenario ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "future"]` to 
`r excreted_total_stats$max[excreted_total_stats$feed == "future"]` g kg$^{-1}$ fish d$^{-1}$ (mean of 
`r excreted_total_stats$mean[excreted_total_stats$feed == "future"]` g kg$^{-1}$ fish d$^{-1}$).

```{r fig-total-excreted-perday}
#| fig-cap: |
#|   Feed left uneaten when fish are fed the reference (blue), past (red) or future (green) feed. Countries are arranged by descending mean latitude.

@@

However, the actual difference in excreted faeces varied by feed, time in the production cycle, and location of the farm [@fig-total-excr-act_diff].

@@

### Composition

```{r excreted composition}
#| eval: true

excreted_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_excr_stat"]]}) %>% bind_rows()
excreted_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_excr_stat"]]}) %>% bind_rows()
excreted_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_excr_stat"]]}) %>% bind_rows()

excreted_all <- excreted_total %>% 
  select(-contains("biom")) %>% 
  merge(excreted_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd"))) %>% 
  rename(protein = mean.x, lipids = mean.y, carbohydrates = mean) %>% 
  mutate(protein = protein/excr_mean %>% set_units("d-1"),
         lipids = lipids/excr_mean %>% set_units("d-1"),
         carbohydrates = carbohydrates/excr_mean %>% set_units("d-1")) %>% 
  select(-excr_mean)

ec <- excreted_all %>% 
  group_by(feed) %>% 
  reframe(P = mean(protein),
          C = mean(carbohydrates),
          L = mean(lipids)) %>% 
  mutate(source = "excreted") %>% 
  rbind(data.frame(
    feed = c("reference", "past", "future"),
    P = c(tP_1, tP_2, tP_3),
    C = c(tC_1, tC_2, tC_3),
    L = c(tL_1, tL_2, tL_3),
    source = rep("feed", 3)
  )) %>% 
  pivot_longer(cols = c(P,L,C), names_to = "macro", values_to = "value", 
               names_transform = list(macro = as.factor), values_transform = list(value = as.numeric)) %>% 
  mutate(source = factor(source, levels = c("feed", "excreted")),
         macro = factor(macro, levels = c("P", "L", "C"), labels = c("Protein", "Lipids", "Carbohydrates")),
         feed = factor(feed, levels = c("reference", "past", "future")),
         source = factor(source, levels = c("feed", "excreted"), labels = c("Feed", "Faeces")))
```

Unlike the uneaten feed, the composition of excreted faeces differs from that of the incoming feed.
[Note that composition does not change through time or with increasing biomass (because digestibility does not change with fish size or temeprature)]{.aside}

```{r composition-throughtime, class.chunk = "aside"}
#| eval: false

excreted_all %>% 
  pivot_longer(names_to = "macro", values_to = "perc", cols = c("protein", "lipids", "carbohydrates")) %>% 
  mutate(macro = factor(macro, levels = c("protein", "lipids", "carbohydrates"))) %>% 
  group_by(farm_ID, feed, macro) %>% 
  reframe(sd = sd(perc) %>% set_units("d-1"),
          mean = mean(perc)) %>% 
  ggplot(aes(x = farm_ID, colour = feed, fill = feed, y = 100*mean, ymin = 100*(mean-sd), ymax = 100*(mean+sd))) +
  geom_point(size = 2) +
  geom_errorbar(width = 0.5) +
  scale_color_manual(values = feed_pal) +
  scale_fill_manual(values = feed_pal) +
  facet_grid(rows = vars(macro)) +
  theme_classic() +
  labs(x = "Macronutrient", y = "Mean composition of faeces")
```

```{r fig-composition-comparison}
#| eval: true
#| fig-cap: Comparison of the composition of the reference (blue), past (red) and future (green) feed and faeces of salmon fed each feed.

@@

## Total nutrients

>>>>

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate", "worrms", "aquamapsdata") %>% sort()

@@

* Hazard: Nutrient input. Present as a change between feeds (%?), relative to background levels (ratio?)
  * Possibly ask Camille how the quotas for Tasmanian salmon are calculated
  * Background levels are from [oceanhealthindex.org](https://knb.ecoinformatics.org/view/resource_map_doi:10.5063/F1610XPS) - double-check that it includes wastewater input AND agriculture - should include both - if they're not combined ask Ben because he has them combined somewhere
  * If this is too uncertain perhaps combine into categorical (oligotrophic/eutrophic) classification?
* Exposure: Species presence. Use IUCN/Aquamap data to estimate the % of each species range is occupied by farms
  * give each farm a 1km radius, check for overlaps
  * Gage already has the data to process the Aquamap data
* Vulnerability: species vulnerability to eutrophication + nutrient input [@butt_trait-based_2022; @ohara_code_2021]. Possibly also use IUCN threat levels for prioritisation?

@@

```{r inputs stats through time}
#| eval: false
#| code-summary: Show how inputs (mass and N) change through time (across all farms)

mass_inputs_ts
N_inputs_ts

# Through time
inputs_stats_ts <- inputs_df %>%
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

inputs_stats_rel_ts <- inputs_rel_df %>% 
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

# Total (at production end)
inputs_stats_end <- inputs_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

inputs_stats_rel_end <- inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

inputs_stats_rel_end_diff <- inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(future = (future-reference)/reference,
         past = (past-reference)/reference) %>% 
  select(-reference) %>% 
  pivot_longer(cols = c(past, future), names_to = "feed", values_to = "diff") %>% 
  mutate(feed = factor(feed, levels = c("past", "future"))) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(diff),
          max = maxna(diff),
          sd = sdna(diff),
          mean = meanna(diff))

# Pull key values (uneaten feed)
u_stmin1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(min)
u_stmax1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(max)
u_stmean1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(mean)
u_stmin2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(min)
u_stmax2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(max)
u_stmean2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(mean)
u_stmin3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(min)
u_stmax3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(max)
u_stmean3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(mean)

# Convert past and future values to % change values
u_stmin2 <- (100*u_stmin2) %>% round(2)
u_stmax2 <- (100*u_stmax2) %>% round(2)
u_stmean2 <- (100*u_stmean2) %>% round(2)
u_stmin3 <- (100*u_stmin3) %>% round(2)
u_stmax3 <- (100*u_stmax3) %>% round(2)
u_stmean3 <- (100*u_stmean3) %>% round(2)

u_stmin1 <- u_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmax1 <- u_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmean1 <- u_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)

# Pull key values (excreted faeces)
e_stmin1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(min)
e_stmax1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(max)
e_stmean1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(mean)
e_stmin2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(min)
e_stmax2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(max)
e_stmean2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(mean)
e_stmin3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(min)
e_stmax3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(max)
e_stmean3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(mean)

# Convert past and future values to % change values
e_stmin2 <- (100*e_stmin2) %>% round(2)
e_stmax2 <- (100*e_stmax2) %>% round(2)
e_stmean2 <- (100*e_stmean2) %>% round(2)
e_stmin3 <- (100*e_stmin3) %>% round(2)
e_stmax3 <- (100*e_stmax3) %>% round(2)
e_stmean3 <- (100*e_stmean3) %>% round(2)

e_stmin1 <- e_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
e_stmax1 <- e_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
e_stmean1 <- e_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
```

```{r waste inputs stats total}
#| code-summary: Show how waste inputs change in total (at the end of a production period) across all farms

mass_inputs_sep <- mass_inputs_end %>% 
  mutate(waste_per_biomass = set_units(total/biomass_total, "g kg_fish-1")) %>% 
  dplyr::select(-c(total, biomass_total)) %>% 
  pivot_wider(names_from = measure, values_from = waste_per_biomass) %>% 
  mutate(total = excr+uneat) %>% 
  pivot_longer(names_to = "measure", values_to = "waste_per_biomass", cols = c(total, excr, uneat), names_transform = list(measure = as.factor)) %>% 
  group_by(feed, measure) %>% 
  reframe(mean_wpb = meanna(waste_per_biomass),
          mean_wpb = sdna(waste_per_biomass) %>% set_units("g kg_fish-1"),
          min_wpb = minna(waste_per_biomass),
          max_wpb = maxna(waste_per_biomass))

twi_min <- mass_inputs_sep$min_wpb[mass_inputs_sep$measure == "total"] %>% round(2)
twi_max <- mass_inputs_sep$max_wpb[mass_inputs_sep$measure == "total"] %>% round(2)
twi_mean <- mass_inputs_sep$mean_wpb[mass_inputs_sep$measure == "total"] %>% round(2)

# There were only very small differences in total uneaten feed between the different feeds, with the past feed increasing uneaten feed by `r u_stmin2`-`r u_stmax2`%  (mean of `r u_stmean2`%) and the future feed increasing uneaten feed by `r u_stmin3`-`r u_stmax3`%  (mean of `r u_stmean3`%).

# * Total excreted faeces across a cohort fed the reference feed ranged from `r e_stmin1` to `r e_stmax1` g kg$^{-1}$ fish, with a mean of `r e_stmean1` g kg$^{-1}$ fish. 
# * Past feed decreased excreted faeces by `r -1*e_stmin2`-`r -1*e_stmax2`%  (mean of `r -1*e_stmean2`%) 
# * Future feed increased excreted faeces by `r e_stmin3`-`r e_stmax3`%  (mean of `r e_stmean3`%)
```

* Total waste inputs across a single production period of fish fed the reference feed ranged from `r twi_min[1]` to `r twi_max[1]` g kg$^{-1}$ fish, with a mean of `r twi_mean[1]` g kg$^{-1}$ fish.
* Total waste inputs across a single production period of fish fed the past feed ranged from `r twi_min[2]` to `r twi_max[2]` g kg$^{-1}$ fish, with a mean of `r twi_mean[2]` g kg$^{-1}$ fish.
* Total waste inputs across a single production period of fish fed the future feed ranged from `r twi_min[3]` to `r twi_max[3]` g kg$^{-1}$ fish, with a mean of `r twi_mean[3]` g kg$^{-1}$ fish.

```{r fig-waste-inputs-total}
#| fig-cap: | 
#|   Total waste intputs (uneaten feed + excreted faeces) across the whole production period at each farm location.
#| fig-width: 9.5

mass_inputs_sep_geometry <- mass_inputs_end %>% 
  mutate(waste_per_biomass = set_units(total/biomass_total, "g kg_fish-1")) %>% 
  dplyr::select(-c(total, biomass_total)) %>% 
  pivot_wider(names_from = measure, values_from = waste_per_biomass) %>% 
  mutate(total = excr+uneat) %>% 
  pivot_longer(names_to = "measure", values_to = "waste_per_biomass", cols = c(total, excr, uneat), names_transform = list(measure = as.factor)) %>% 
  group_by(farm_ID, feed, measure) %>% 
  reframe(mean_wpb = meanna(waste_per_biomass),
          sd_wpb = sdna(waste_per_biomass) %>% set_units("g kg_fish-1"),
          min_wpb = minna(waste_per_biomass),
          max_wpb = maxna(waste_per_biomass)) %>% 
  merge(farms_geometry, by = "farm_ID")

mass_inputs_robinson <- mass_inputs_sep_geometry %>% 
  filter(feed == "reference" & measure == "total") %>% 
  mutate(mean_wpb = mean_wpb %>% drop_units()) %>% 
  st_sf() %>% 
  st_transform(crs = "+proj=robin")

ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = mass_inputs_robinson, aes(color = mean_wpb), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = expression("Mean waste inputs (g g fish"^-1*")"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()
```

```{r patchwork-waste-map-mercator}
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

mass_inputs_patch <- mass_inputs_sep_geometry %>% 
  filter(feed == "reference" & measure == "total") %>% 
  mutate(mean_wpb = mean_wpb %>% drop_units())

p1 <- ggplot() +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = mass_inputs_patch, aes(color = mean_wpb, geometry = geometry), size = 2) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = expression("Mean waste inputs (g g fish"^-1*")"),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1.5,
    barwidth = 1.5, barheight = 20,
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

```{r fig-total-excreted-perday}
#| eval: false
#| fig-cap: |
#|   Feed left uneaten when fish are fed the reference (blue), past (red) or future (green) feed. Countries are arranged by descending mean latitude.

@@

@@

## Total nutrient inputs

```{r protein raw data}
#| code-summary: Load and prep macronutrients data (keep in g faeces/feed)

N_inputs_df <- rbind(
  farmrun_files %>% 
    purrr::map(function(f) {qs::qread(f)[["P_excr_stat"]]}) %>% 
    bind_rows() %>% 
    mutate(measure = "excr"),
  farmrun_files %>% 
    purrr::map(function(f) {qs::qread(f)[["P_uneat_stat"]]}) %>% 
    bind_rows() %>% 
    mutate(measure = "uneat")
) %>% 
  mutate(measure = as.factor(measure),
         t = as.integer(t))

# Absolute inputs
N_inputs_df <- N_inputs_df %>% 
  mutate(mean = mean %>% set_units("g") %>% set_units("gN"),
         sd = sd %>% set_units("g") %>% set_units("gN"))

# Inputs relative to biomass produced
N_inputs_rel_df <- N_inputs_df %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(mean = mean.x, 
         sd = sd.x, 
         biomass_mean = mean.y,
         biomass_sd = sd.y) %>% 
  mutate(biomass_mean = biomass_mean %>% set_units("g_fish"),
         biomass_sd = biomass_sd %>% set_units("g_fish"),
         biomass_sd = biomass_sd/biomass_mean,
         sd = sd/mean,
         mean = mean/biomass_mean,
         sd = sqrt(sd^2 + biomass_sd^2) * mean,
         biomass_sd = biomass_sd * biomass_mean)
```

```{r fig-composition-comparison}
#| eval: false
#| fig-cap: Comparison of the composition of the reference (blue), past (red) and future (green) feed and faeces of salmon fed each feed.

@@

```{r composition stats}
# Through time
N_inputs_ts <- N_inputs_df %>%
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

N_inputs_rel_ts <- N_inputs_rel_df %>% 
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

# Total (at production end)
N_inputs_end <- N_inputs_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

N_inputs_rel_end <- N_inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

N_inputs_rel_end_diff <- N_inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(future = (future-reference)/reference,
         past = (past-reference)/reference) %>% 
  select(-reference) %>% 
  pivot_longer(cols = c(past, future), names_to = "feed", values_to = "diff") %>% 
  mutate(feed = factor(feed, levels = c("past", "future"))) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(diff),
          max = maxna(diff),
          sd = sdna(diff),
          mean = meanna(diff))

# Pull key values (uneaten feed)
u_stmin1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(min)
u_stmax1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(max)
u_stmean1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(mean)
u_stmin2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(min)
u_stmax2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(max)
u_stmean2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(mean)
u_stmin3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(min)
u_stmax3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(max)
u_stmean3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(mean)

# Convert past and future values to % change values
u_stmin2 <- (100*u_stmin2) %>% round(2)
u_stmax2 <- (100*u_stmax2) %>% round(2)
u_stmean2 <- (100*u_stmean2) %>% round(2)
u_stmin3 <- (100*u_stmin3) %>% round(2)
u_stmax3 <- (100*u_stmax3) %>% round(2)
u_stmean3 <- (100*u_stmean3) %>% round(2)

u_stmin1 <- u_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmax1 <- u_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmean1 <- u_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
```
